---
title: "Feature and Target Engineering"
output:
  rmdformats::downcute:
    self_contained: true
    downcute_theme: "chaos"
    thumbnails: false
    lightbox: true
    gallery: false
    code_folding: show
    toc_depth: 3
---


```{r setup, include=FALSE}
## Global options
knitr::opts_chunk$set(cache = TRUE, warning = FALSE)
```


```{r prerequisites, message=FALSE}

library(dplyr)
library(ggplot2)
library(visdat)

library(caret)
library(recipes)

# ames housing data
ames <- AmesHousing::make_ames()

set.seed(123)

# stratified sampling with the rsample package
split <- rsample::initial_split(ames, prop = 0.7, strata = "Sale_Price")
ames_train <- rsample::training(split)
ames_test <- rsample::testing(split)
```

## Target Engineering

There are two main approaches to help correct for positively skewed target variables.

**1. Normalize with log-transformation**

```{r}
ames_recipe <- recipe(Sale_Price ~ ., data = ames_train) %>% 
  step_log(all_outcomes())

ames_recipe
```

If the response has negative values or zeros then log transform will produce `NaNs` and `-Infs` respectively. If the nonpositive response values are small (in between -0.99 to 0), then using `log1p()` (which adds 1 prior to applying log transform) could be an option. (similarly there's an `offset` argument for `step_log`)

```{r}
recipe(Sale_Price ~ ., data = ames_train) %>% 
  step_log(all_outcomes(), offset = 1)
```

But if the data contains neg values $\le -1$, we need to use Yeo-Johnson transformation.

**2. Use a Box-Cox transformation**

Its more flexible than log-transformation. Yeo-Johnson transformation is very similar to Box-cox, but doesn't require the input variables to be strictly positive.

```{r}
ames_recipe <- recipe(Sale_Price ~ ., data = ames_train) %>% 
  step_BoxCox(all_outcomes())
```


```{r}
recipe(Sale_Price ~ ., data = ames_train) %>% 
  step_YeoJohnson(all_outcomes())
```

## Dealing with Missingness

Data can missing for many different reasons.

- **Informative missingness**: Implies a structural cause for data missing and happens due to defficiencies how the data was collected or due to abnormalities in the observational environment. So, such missingness can provide some insight. we can label such missing values as a unique category (such as, `"None"`).

- **Missingness at random (MAR)**: missing values occur independently of the data collection process and depending on the data-size we may impute or delete them.

- **GLM models, neural network, svm cannot handles missingness**
- **Tree based models have built-in procedure to handle missingness**

But its better to handle missingness before the modelling process.


### Visualizing missing values

```{r}
sum(is.na(AmesHousing::ames_raw))
```

We can use heatmaps to visualize the $dist^n$ of missing values for small to medium sized data.


```{r, fig.height=10}
AmesHousing::ames_raw %>% 
  is.na() %>% 
  reshape2::melt() %>% 
  ggplot(aes(Var2, Var1, fill = value)) +
  geom_raster() +
  coord_flip() + 
  scale_y_continuous(NULL, expand = c(0, 0)) +
  scale_fill_grey(name = "", 
                  labels = c("Present", "Missing")) + 
  theme(axis.text.y = element_text(size = 10))
```

```{r}
AmesHousing::ames_raw %>% 
  filter(`Garage Cars` == 0 & `Garage Area` == 0) %>% 
  select(starts_with("Garage"))
```

we see that for houses whose have no Garage Cars and Garage Area, all related variables about Garage have missing values. So this missing is informative. All `NA` values for these variables may be imputed with a `"None"` category.


```{r, fig.height=6, fig.width=12}
visdat::vis_miss(AmesHousing::ames_raw, cluster = TRUE) + 
  theme(
    axis.text.x.top = element_text(size = 8, angle = 90)
  )
```


### Imputation

Imputation is the process of replacing the missing values with the best guessed value.

We can perform imputation based on mean, median or mode (for categorical) of a feature. Also as an alternative, we can do model based imputation (KNN or tree based imputation).

#### Estimated Statistic based imputation

```{r}
ames_recipe  %>% 
  step_impute_median(Gr_Liv_Area)

```

#### KNN based imputation

KNN imputes missing values by identifying obsn with missing values at first and then identify other similar (based on other feature) obsns and use the values from these nearest neighbor to impute missing values.

In this method, missing value for a given obsn is treated as response and predicted by the avg (for quantitative values) or mode (for categorical values) of the k-nearest neighbors.

For small to medium sized data, knn imputation can be used. For large data, its better avoided.


```{r}
ames_recipe %>% 
  step_impute_knn(all_predictors(), neighbors = 6)
```

#### Bagged tree based imputation

Bagged trees offer a compromise between predictive and computational burden.

It works similar as knn (ie. identify missing val as response)


```{r}
ames_recipe %>% 
  step_impute_bag(all_predictors())
```


## Feature Filtering

Features with zero variance and near zero variance should be removed. Because they provide no useful information for the model.

we can detect them in the following way

```{r}
caret::nearZeroVar(ames_train, saveMetrics = TRUE) %>% 
  tibble::rownames_to_column() %>% 
  filter(nzv) %>% 
  DT::datatable()

```


```{r}
ames_recipe %>% 
  step_nzv(all_nominal())
```

## Numeric feature Engineering

### handling skewness

Models that have parametric distributional assumption can greatly affected by the skewness of numeric feature. So to minimize the skewness we can consider either box-cox transformation or yeo-johnson transformation.


```{r}
recipe(Sale_Price ~ ., data = ames_train) %>% 
  step_YeoJohnson(all_numeric())
```

### standardization

```{r}
ames_recipe %>% 
  step_center(all_numeric(), -all_outcomes()) %>% 
  step_scale(all_numeric(), -all_outcomes())
```

## Categorical feature engineering


### lumping (Reducing categories)

We can reduce the categories with few observations. Also for numeric feature, binning seems sometime very useful

```{r}
count(ames_train, Neighborhood) %>% arrange(n)

count(ames_train, Screen_Porch) %>% arrange(n)
```

However,

- Lumping should be done sparingly, as this sometimes may lead to loss in model performance

- **Tree based models often perform exceptionally well with high cardinality features and are not impacted by levels with few obsn**

```{r}
# lump levels for two features
lumping <- recipe(Sale_Price ~ ., data = ames_train) %>% 
  step_other(Neighborhood, threshold = 0.01, other = "other") %>% 
  step_other(Screen_Porch, threshold = 0.1, other = ">0")


# preping and baking
apply_2_training <- prep(lumping, training = ames_train) %>% 
  bake(ames_train)
```


```{r}
count(apply_2_training, Neighborhood) %>% arrange(n)
```

```{r}
count(apply_2_training, Screen_Porch) %>% arrange(n)
```


### One-hot and dummy encoding

In one hot encoding we create separate variable for each level of a categorical variable. This is called **less than full rank encoding**, but this creates perfect multicollinearity which caused problems in some predictive modelling algorithm like OLS, neural net.

Alternatively, we can do dummy encoding, where we drop one of the levels and create one less variable than the number of categorical levels. This is **full-rank encoding**.


```{r}
recipe(Sale_Price ~ ., data = ames_train) %>% 
  step_dummy(all_nominal(), one_hot = TRUE)
```

### Label encoding and Ordinal encoding

Label encoding is the process of turning the categorical levels into integers. Now if the categorical variable is a factor with prespecified levels then numeric conversion will be in level order, otherwise it will be in alphabetic order.

```{r}
# original categories
ames_train %>% count(MS_SubClass)

# label encoded
recipe(Sale_Price ~ . , data = ames_train) %>% 
  step_integer(MS_SubClass) %>% 
  prep(ames_train) %>% 
  bake(ames_train) %>% 
  count(MS_SubClass)
```

Now we should be careful about label encoding unordered categorical features because most models will treat them as ordered numeric features. But label encoding is a natural choice for ordered factors, in which case it is known as Ordinal Encoding.


```{r}
# original category with natural order
ames_train %>% count(Overall_Qual)

# ordinal encoding
recipe(Sale_Price ~ ., data = ames_train) %>% 
  step_integer(Overall_Qual) %>% 
  prep(ames_train) %>% 
  bake(ames_train) %>% 
  count(Overall_Qual)
```


## Dimension reduction

Its an alternative approach to automatically filter out non-informative approach.

```{r}
recipe(Sale_Price ~ ., data = ames_train) %>% 
  step_center(all_numeric()) %>% 
  step_scale(all_numeric()) %>% 
  step_pca(all_numeric(), threshold = 0.95)

```

